Circular Regression Trees and Forests with an Application to Probabilistic Wind Direction Forecasting

### 2020-07-06 pkg/circtree/inst/wind_paper/draft/JRSSC/rebuttal.tex

We first want to thank you for the positive feedback and your valuable input.

We agree with your comment that additional information on the calculation
requirements is valuable for the user. We have therefore listed the needed
computing time of the different methods on a standard desktop computer within
the section 'Computational details'.

\vspace{0.5em}
\begin{re}
Is it possible to include a link to computer code to reproduce your
analysis? This is not essential for publication but would be helpful for
readers who wish to perform similar analyses. The R package you have provided
for your method is very helpful in this regard.  
\end{re}

We have used data from the Austrian air navigation service provider, which
we are not at liberty to share. Hence, we are unfortunately unable to provide
code and data to reproduce the application. However, as you have already
pointed out the implementation of circular regression trees and forests is
provided in a freely available R package with examples included in the documentation.
Independent of this publication, we are planning to provide an updated online
platform for the whole `partykit' project in the near future.

### 2020-07-06 pkg/circtree/inst/wind_paper/draft/JRSSC/rebuttal.tex

First of all we want to thank you for your positive feedback and valuable input.

Indeed, information criteria like AIC or BIC are natural candidates for
determining the model order in distributional models, i.e., the number of splits
in distributional trees. While they could be used instead of the $p$-values in
the ``internal stopping'' or ``pre-pruning'', they are typically considered
for ``post-pruning'' after initially growing a large tree (e.g., in
Su~\emph{et al.}~2004 or Zeileis~\emph{et al.}~2008). This is now discussed
in Appendix~A along with the other details of the tree algorithm employed.
There we also refer to the recent work of Schlosser~\emph{et al.}~2019
which empirically investigates the performance of different pre- and post-pruning
strategies in unbiased recursive partitioning. This work finds that post-pruning
only improves the performance of the unbiased trees considered when the
underlying tests do not work well, e.g., when the power of the tests is low
or the sample size extremely large.

\vspace{0.5em}
\begin{re}
I would like to point out related works on circular regression models for
shape representation (Kashyap and Chellappa) and circular random forests for
vehicle direction estimation (Hara and Chellappa) given below. I feel there is
some overlap between this paper and the work of Hara and Chellappa on using
circular statistics for vehicle orientation estimation. The authors may include
some comparisons with Hara, Chellappa work.
\end{re}

Thank you very much for pointing out these interesting references that, unfortunately,
we were not aware of before. We now briefly discuss the work of Hara and Chellappa~(2017),
comparing their approach to ours in Sect.~1.1. The main difference is that their
algorithm aims at conditional mean estimation for circular data and only adopts
distributional aspects for determining the number of splits. Specifically, the
variances are restricted to be shared across splits whereas in our proposal they
are full model parameters that are estimated and tested along with the means.

@@

Hara K, Chellappa R (2017). ``Growing Regression Tree Forests by
Classification for Continuous Object Pose Estimation.'' \emph{International
Journal of Computer Vision}, \textbf{122}\,(2), 292--312.
doi:10.1007/s11263-016-0942-1.

Kashyap R, Chellappa R (1981). ``Stochastic Models for Closed Boundary
Analysis: Representation and Reconstruction.'' \emph{IEEE Transactions on
Information Theory}, \textbf{27}\,(5), 627--637. doi:10.1109/TIT.1981.1056390.

Schlosser L, Hothorn T, Zeileis A (2019). ``The Power of Unbiased Recursive
Partitioning: A Unifying View of CTree, MOB, and GUIDE.'' arXiv:1906.10179,
arXiv.org E-Print Archive.

Su X, Wang M, Fan J (2004). ``Maximum Likelihood Regression Trees.'' 
\emph{Journal of Computational and Graphical Statistics}, \textbf{13}\,(3), 586--598. 
doi:10.1198/106186004X2165.

### 2020-07-06 pkg/circtree/inst/wind_paper/draft/draft.Rnw
Minor adaptions and spell checking
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@3051 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 8adb6cdd8382022417f7a830ad58c451cdf6a467

This novel approach to circular regression trees and forests complements the
literature on tree-based circular modeling. \cite{Lund:2002} already introduced
a circular regression tree algorithm where binary splits are made based on an
angular distance measure capturing node homogeneity. However, this only models
changes in the conditional mean, whereas modeling the conditional variance or
full probabilistic distribution would also allow uncertainty in the forecast to
be estimated \citep{Gneiting:2008}.  Moreover, \citet{Hara+Chellappa:2017}
introduced both regression trees and random forests for image-based object
direction estimation. These extend classical conditional-mean regression trees
\citep{Breiman+Friedman+Stone:1984} and corresponding random forests
\citep{Breiman:2001} to conditional mean models for cicular data. But rather
than considering only binary splits, they employed $k$-means clustering
to determine multi-way splits with the number $k$ adaptivaly determined
based on the Bayes information criterion with either a Gaussian or von~Mises
kernel. While a distributional aspect is considered in this step of the
algorithm, variances are always restricted to be shared across splits and, more
importantly, no probabilistic forecasts are obtained. Therefore, our proposed
circular regression trees and random forests take these ideas a step further:
A fully distributional approach is adopted for all aspects of the algorithms
from selection of the split variables, selection of the split points, and
ensemble aggregation of the predictions across trees in a forest. Therefore,
this yields a full probabilistic predictive model as proposed by
\citet{Gneiting:2008}.

>>>>

This novel approach to circular regression trees and forests complements the
literature on tree-based circular modeling. \cite{Lund:2002} already introduced
a circular regression tree algorithm where binary splits are made based on an
angular distance measure capturing node homogeneity. However, this only models
changes in the conditional mean, whereas modeling the conditional variance or
full probabilistic distribution would also allow uncertainty in the forecast to
be estimated \citep{Gneiting:2008}.  Moreover, \citet{Hara+Chellappa:2017}
introduced both regression trees and random forests for image-based object
direction estimation. These extend classical conditional-mean regression trees
\citep{Breiman+Friedman+Stone:1984} and corresponding random forests
\citep{Breiman:2001} to conditional mean models for circular data. But rather
than considering only binary splits, they employed $k$-means clustering
to determine multi-way splits with the number $k$ adaptively determined
based on the Bayes information criterion with either a Gaussian or von~Mises
kernel. While a distributional aspect is considered in this step of the
algorithm, variances are always restricted to be shared across splits and, more
importantly, no probabilistic forecasts are obtained. Therefore, our proposed
circular regression trees and random forests take these ideas a step further:
A fully distributional approach is adopted for all aspects of the algorithms
from selection of the split variables, selection of the split points, and
ensemble aggregation of the predictions across trees in a forest. This provides
a full probabilistic predictive model, as proposed by \citet{Gneiting:2008}.
